---
title: 'Two-stage exams: Study 2'
author: "George Kinnear"
date: "20/06/2020"
output:
  pdf_document:
    toc: true
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.path='rmd_figs_study2/',
                      warning=FALSE, message=FALSE)

library(tidyverse)
library(ggthemes)
library(ggpubr)
library(gt)        # For formatting tables


theme_set(
  theme_pubr() +
    theme(legend.position = "right")
)
heathers = c("#00b3c3","#e51c24","#016637","#fcb123")    #original green: #008a49

```

# Data

Import and combine the datasets.

```{r data}

data = read.csv('Study2_data.csv', header=T) %>%
  dplyr::select(-c(X)) %>%
  mutate(Q = fct_relevel(Q, "Q10", after=Inf))


S1234data = subset(data,
                   select=c('Q','Student','Stage1score','Stage2score','Stage3score','Stage4score')) %>%
  filter(!is.na(Stage2score))

S123data = subset(data,select=c('Q','Student','Stage1score','Stage2score','Stage3score','ZippGroup'))
S123data = S123data[complete.cases(S123data),]
```


## View a summary of the data

```{r data-check}
ld <- gather(data = subset(data,select=c('Q','Student','Stage1score','Stage2score','Stage3score')), 
             key = stage, 
             value = score, 
             Stage1score, Stage2score, Stage3score)
ld <- ld[complete.cases(ld),]

# Code from here:
# https://www.r-bloggers.com/building-barplots-with-error-bars/

plotData <- aggregate(ld$score,
                      by = list(Q = ld$Q, Stage = ld$stage),
                      FUN = function(x) c(mean = mean(x), sd = sd(x),
                                          n = length(x)))
plotData <- do.call(data.frame, plotData)
plotData$se <- plotData$x.sd / sqrt(plotData$x.n)

colnames(plotData) <- c("Q", "Stage", "mean", "sd", "n", "se")

limits <- aes(ymax = plotData$mean + plotData$se,
              ymin = plotData$mean - plotData$se)

p <- ggplot(data = plotData, aes(x = factor(Q), y = mean,
                                 fill = factor(Stage))) +
  geom_bar(stat = "identity",
           position = position_dodge(0.9)) +
  labs(x = "Question",
       y = "Percentage of students answering correctly",
       fill = "Stage") +
  scale_fill_manual(values=heathers) +
  scale_y_continuous(labels = scales::percent)
p
```


## Mean at each stage (as bars)
```{r fig-means-bars}
ld <- gather(data = S1234data, 
             key = stage, 
             value = score, 
             Stage1score, Stage2score, Stage3score, Stage4score)
ld <- ld[complete.cases(ld),]

plotData <- aggregate(ld$score,
                      by = list(Q = ld$Q, Stage = ld$stage),
                      FUN = function(x) c(mean = mean(x), sd = sd(x),
                                          n = length(x)))
plotData <- do.call(data.frame, plotData)
plotData$se <- plotData$x.sd / sqrt(plotData$x.n)

colnames(plotData) <- c("Q", "Stage", "mean", "sd", "n", "se")

plotData = plotData %>%
  mutate(
    Stage = gsub(".*(\\d).*","\\1",Stage)  # alternatively: Stage = parse_number(Stage)
  ) %>%
  arrange(Q,Stage)
  
plotCounts = plotData %>% group_by(Q,Stage) %>% dplyr::select(Stage,n) %>%
  mutate(scale_lab = paste0("Stage ",Stage, " (n=",n,")")) %>%
  ungroup

limits <- aes(ymax = plotData$mean + plotData$se,
              ymin = plotData$mean - plotData$se)

ggplot(data = plotData, aes(x = factor(Q), y = mean,
                                 fill = factor(Stage), label=n)) +
  geom_bar(stat = "identity",
             position = position_dodge(0.9))+
  geom_text(position = position_dodge(width = 0.9),aes(y=-0.05),angle=90) +
  geom_errorbar(limits, position = position_dodge(0.9),
                width = 0.25)  +
  labs(x = "Question",
       y = "Percentage of students answering correctly",
       fill = "Stage") +
  scale_fill_manual(values=heathers, labels=paste0("Stage ",c(1:4))) + #plotCounts$scale_lab) +
  #scale_fill_grey() +
  scale_y_continuous(labels = scales::percent)
ggsave("Figs/Study2_S1234_means.pdf",width=20,height=10,units="cm",dpi=300)
```

## Mean at each stage (as points)
```{r fig-means-pts}
ggplot(data=plotData,aes(x=Stage,y=mean,group=Q,label=n))+
  geom_line()+
  geom_errorbar(aes(ymax = plotData$mean + plotData$se,
                    ymin = plotData$mean - plotData$se),
                position = position_dodge(0.9),
                width = 0.5)  +
  geom_point(position = position_dodge(0.9),size=5,aes(color=Stage))+
  scale_color_manual(values=heathers, labels=paste0("Stage ",c(1:4),rep(c(""," (groups)"),2))) +
  labs(x = "Stage",
       y = "Percentage of students answering correctly",
       fill = "Stage") +
  scale_y_continuous(labels = scales::percent)+
  facet_grid(cols=vars(Q))+
  coord_cartesian(ylim=c(0,1),clip="off")+
  geom_text(position = position_dodge(width = 0.9),
            aes(y=-0.01, label=paste0("",plotCounts$n)),
            angle=90,
            color="#777777") +
  theme(strip.background = element_rect(fill=NA,colour = NA),
        strip.text = element_text(size=12, face="bold"))
ggsave("Figs/Study2_S1234_means_pts.pdf",width=20,height=10,units="cm",dpi=300)
```

## Mean at each stage (table, with standard errors)

```{r tab-means}
tab = plotData %>% 
  mutate(
    entry = paste0(sprintf("%2.0f", mean*100), " (", sprintf("%2.1f", se*100), ")"),
    Stage = gsub(".*(\\d).*","\\1",Stage)
  ) %>%
  group_by(Q,Stage) %>%
  select(Q,Stage,entry) %>%
  spread(Q,entry)

tab %>% knitr::kable()
```


# Forming the Zipp tables

This constructs the data in Table 4 of the paper.

## Stages 1-3 only
```{r zip-tab-123}
zipptab = S123data %>%
  group_by(ZippGroup) %>%
  summarize( numcorrect = sum(Stage3score),
             numingroup = n()) %>%
  mutate( pc = numcorrect/numingroup,
          entry = paste0(sprintf("%2.1f", pc*100), " (", numcorrect, "/",numingroup,")"))
print("Entries of the Zipp table for Stage 3 only")
zipptab %>% knitr::kable()
```


# Group dynamics: Stage 1 vs Stage 2

Here we look at the relative performance in the groups across the first two stages.

This is discussed in section 5.2.2 of the paper.

```{r group-correctness}

dataStudents = read.csv('Study2_data_raw_stage1.csv', header=T)
dataGroups = read.csv('Study2_data_raw_stage2.csv', header=T)

responses = dataStudents %>%
  select(AnonKey, contains("Q")) %>%
  gather(key,Response,-AnonKey) %>%
  tidyr::extract(key, c("Week", "Question"), regex="Wk(.*)Q(.*)")

qnkey = responses %>%
  filter(AnonKey=="KEY")

studentGroups = dataStudents %>%
  filter(AnonKey != "KEY") %>%
  select(AnonKey, contains("Gp")) %>%
  gather(key,Group,-AnonKey) %>%
  tidyr::extract(key,"Week",regex="Wk(.*)Gp")

groupResponses = dataGroups %>%
  gather(Question,Response,-Group) %>%
  tidyr::extract(Question,"Question",regex="Q(.*)")

S12data = responses %>%
  filter(AnonKey!="KEY") %>%
  left_join(studentGroups,by=c("AnonKey","Week")) %>%
  rename( Stage1response = Response ) %>%
  left_join(groupResponses,by=c("Group","Question")) %>%
  rename( Stage2response = Response) %>%
  filter(Stage1response!="")

# Check if the ultimate response is correct
S12data_scored = S12data %>%
  left_join( qnkey %>% select(-AnonKey) %>% rename(Key=Response), c("Week","Question")) %>%
  mutate(
    Stage1score = case_when(
      Stage1response=="" ~ NA_real_,
      Stage1response==Key ~ 1,
      Stage1response!=Key ~ 0,
      TRUE ~ NA_real_
    ),
    Stage2score = case_when(
      Stage2response=="" ~ NA_real_,
      Stage2response==Key ~ 1,
      Stage2response!=Key ~ 0,
      TRUE ~ NA_real_
    ),
    Stage2scorePartial = case_when(
      Stage2response=="" | is.na(Stage2response) ~ NA_real_,
      Stage2response==Key ~ 1,
      str_length(Stage2response)==2 & str_sub(Stage2response,2,2)==Key ~ 0.5,
      TRUE ~ 0
    )
  ) %>%
  select(-Key) %>%
  rename(Student=AnonKey)

S1modalresponse = S12data_scored %>%
  group_by(Group,Question,Stage1response) %>%
  tally() %>%
  arrange(Group,Question,-n) %>%
  mutate(
    nextmost = lead(n),
    topmost = max(n),
    ismodal = ifelse(n!=nextmost & n==topmost,1,0)
  ) %>%
  filter(ismodal==1) %>%
  select(Group,Question,Stage1response) %>%
  rename(S1ModalResponse = Stage1response)

groupPerf = S12data_scored %>%
  group_by(Week,Group,Question) %>%
  summarise(
    n=length(Student),
    tot=sum(Stage1score),
    mean=mean(Stage1score),
    Responses = str_c(Stage1response,collapse = "")
  ) %>%
  left_join(S12data_scored %>% select("Group","Question","Stage2response",
                                      "Stage2score","Stage2scorePartial") %>% distinct(),
            by=c("Group","Question")) %>%
  left_join(S1modalresponse) %>%
  ungroup() %>%
  left_join( qnkey %>% select(-AnonKey) %>% rename(Key=Response)) %>%
  mutate(
    Stage2responseFirst = str_sub(Stage2response,1,1),
    WentWithModal = ifelse(S1ModalResponse==Stage2responseFirst,1,0),
    modalCorrect = ifelse(S1ModalResponse==Key,1,0),
    usedModalSecond = ifelse(S1ModalResponse==str_sub(Stage2response,2,2),1,0)
  )

groupPerf %>%
  group_by(WentWithModal)%>%
  summarise(
    count = n(),
    correctness = mean(Stage2score)
  ) %>% knitr::kable()
```

On 110 out of 250 occasions, the group went with the modal response,
and about 84% of the time they were correct to do so
96 times the group did not have a clear modal option.

```{r group-perf-modal}
groupPerf %>%
  group_by(WentWithModal,modalCorrect) %>%
  summarise(
    count = n(),
    correctness = mean(Stage2score)
  ) %>% knitr::kable()
```

When the group did not use the modal response, it was correct 9/44 times.

Did they go on to try this as their next answer?
```{r group-perf-next-ans}
groupPerf %>%
  group_by(WentWithModal,modalCorrect,usedModalSecond) %>%
  summarise(
    count = n(),
    correctness = mean(Stage2score),
    partial = mean(Stage2scorePartial)
  ) %>% knitr::kable()
```

Yes, in 7/9 cases they did

These next tables look at a different way of slicing the data:

```{r group-perf}
groupPerf %>%
  group_by(modalCorrect,WentWithModal) %>%
  summarise(
    count = n(),
    correctness = mean(Stage2score),
    partial = mean(Stage2scorePartial)
  ) %>% knitr::kable()

groupPerf %>%
  group_by(WentWithModal) %>%
  summarise(
    count = n(),
    correctness = mean(Stage2score),
    partial = mean(Stage2scorePartial)
  ) %>% knitr::kable()
```

## Stage 1 vs Stage 2
Exploring the association between individuals at stage 1 and the group correctness at stage 2, plotting this in different ways.

```{r stage1vs2}
S2meanByS1 = groupPerf %>%
  group_by(mean) %>%
  summarise(
    S2mean = mean(Stage2scorePartial)
  )

ggplot(S2meanByS1, aes(x=mean, y=S2mean)) + geom_point()

ggplot(groupPerf, aes(x=mean, y=Stage2scorePartial)) + geom_point(position="jitter")

ggplot(groupPerf, aes(x=mean, y=Stage2scorePartial)) + geom_point() +
  geom_count() + scale_size_area(max_size = 10)

ggplot(groupPerf, aes(x = Stage2scorePartial, y = mean, group=Stage2scorePartial)) + 
  geom_boxplot()+ 
  scale_y_continuous(breaks=seq(0,100,10))

groupPerfS12 = groupPerf %>%
  mutate(
    tot_group = cut(tot,breaks=c(-Inf,0.5,1.5,2.5,Inf),labels=c("0","1","2","3 or more"))
  ) %>%
  group_by(tot_group) %>%
  mutate(
    Stage2scorePartial = ceiling(Stage2scorePartial) # round this up so that groups scoring 0.5 count as 1 correct group
  ) %>%
  summarise(
    S2avg = mean(Stage2score),
    S2se = sd(Stage2score)/sqrt(n()),
    S2n = n(),
    S2Pavg = mean(Stage2scorePartial),
    S2Pse = sd(Stage2scorePartial)/sqrt(n())
  )
  

```

```{r group-perf-plot}

ggplot(groupPerfS12,aes(x=tot_group,y=S2avg,label=S2n))+
  geom_errorbar(aes(ymax = groupPerfS12$S2avg + groupPerfS12$S2se,
                    ymin = groupPerfS12$S2avg - groupPerfS12$S2se),
                position = position_dodge(0.9),
                width = 0.1)+
  geom_point(aes(colour="First attempt"),size=5)+
  geom_errorbar(aes(ymax = groupPerfS12$S2Pavg + groupPerfS12$S2Pse,
                    ymin = groupPerfS12$S2Pavg - groupPerfS12$S2Pse),
                position = position_nudge(x=0.1),
                width = 0.1)+
  geom_point(aes(y=S2Pavg,colour="Second attempt"),position=position_nudge(x=0.1),size=5)+
  scale_y_continuous(labels = scales::percent,breaks=seq(0,1,by=.2))+
  scale_color_manual(values=heathers) +
  coord_cartesian(ylim=c(0,1),clip="off")+
  geom_text(position = position_dodge(width = 0.9),
            aes(y=-0.01, label=paste0("n=",groupPerfS12$S2n)),
            angle=0,
            color="#777777")+
  labs(x = "Number of students correct at stage 1",
       y = "Groups answering correctly",
       colour = "Stage 2 attempt") +
  theme(strip.background = element_rect(fill=NA,colour = NA),
        strip.text = element_text(size=12, face="bold"))
ggsave("Figs/Study2_S12_collab.pdf",width=15,height=7,units="cm",dpi=300)
```


## Group dynamics

This replicates the analysis of Levy et al. (2018), producing Fig 2 of the paper. There is extra detail here, with the various measures like `collaborative efficiency' shown for each group and also plotted.

Find the top scoring student in each group,
and the "super" score (max score across all students in the group, by question)


```{r analyse-stages12}
S1superandtop = S12data_scored %>%
  mutate(
    Week = as.numeric(Week)
  ) %>%
  group_by(Group,Week,Question) %>%
  mutate(
    superstudent = max(Stage1score)
  ) %>%
  group_by(Group,Week,Student) %>%
  mutate(
    topstudent = sum(Stage1score)/n() # the Student's mean score on the n() Questions
  ) %>%
  group_by(Group,Week) %>%
  summarise(
    superstudent = sum(superstudent)/n(),
    topstudent = max(topstudent)
  )

LevyA = S12data_scored %>%
  mutate(
    Week = as.numeric(Week)
  ) %>%
  group_by(Student,Week) %>%
  summarise(
    Stage1pc = sum(Stage1score)/n()
  ) %>%
  group_by(Week) %>%
  summarise(
    S1mean = mean(Stage1pc),
    S1sd = sd(Stage1pc),
    S1n = n()
  )

LevyByGroup = groupPerf %>%
  mutate(
    Week = as.numeric(Week)
  ) %>%
  left_join(S1superandtop) %>%
  left_join(LevyA %>% select(Week,S1sd)) %>%
  arrange(Week) %>%
  group_by(Week,Group) %>%
  summarise(
    n = max(n),
    sd = max(S1sd),
    IndivA = mean(mean),
    GroupB = mean(Stage2scorePartial),
    TopC = max(topstudent),
    SuperD = max(superstudent),
    GainBA = (GroupB-IndivA)/sd,
    TopSurplus = (TopC-IndivA)/sd,
    SuperSurplus = (SuperD-IndivA)/sd,
    CollabEfficiency = GainBA / SuperSurplus
  ) %>%
  ungroup()
LevyByGroup %>% knitr::kable(digits = 2)
```

```{r levy-plots}

ggplot(stack(LevyByGroup %>% select(IndivA,GroupB,TopC,SuperD)), aes(x = ind, y = values)) +
  geom_violin(fill=heathers[1]) +
  geom_boxplot(width=0.2,color=heathers[2],lwd=1) +
  geom_jitter(shape=16, position=position_jitter(0.2),alpha=0.5) +
  labs(x = "Average score of...",
       y = "Percentage correct") +
  scale_y_continuous(labels = scales::percent)
ggsave("Figs/Study2_LevyABCD.pdf",width=20,height=10,units="cm",dpi=300)
ggsave("Figs/Study2_LevyABCD_small.pdf",width=10,height=7,units="cm",dpi=300)

ggplot(stack(LevyByGroup %>% select(GainBA,TopSurplus,SuperSurplus,CollabEfficiency)), aes(x = ind, y = values)) +
  geom_violin(fill=heathers[1]) +
  geom_boxplot(width=0.2,color=heathers[2],lwd=1) +
  geom_jitter(shape=16, position=position_jitter(0.2),alpha=0.5) + 
  labs(x = "Difference in average scores",
       y = "Difference (in SDs)")+
  theme(axis.text.x = element_text(angle = 15, hjust = 1))
ggsave("Figs/Study2_LevyDiffs.pdf",width=20,height=10,units="cm",dpi=300)
ggsave("Figs/Study2_LevyDiffs_small.pdf",width=10,height=7,units="cm",dpi=300)

```

```{r levy-summary}

LevyByGroup %>%
  summarise(
    CollabEfficiency_m = mean(CollabEfficiency),
    CollabEfficiency_sd = sd(CollabEfficiency),
    n = n()
  )
LevyByGroup %>%
  filter(CollabEfficiency>1) %>%
  count() %>%
  knitr::kable()
LevyByGroup %>%
  filter(CollabEfficiency==1) %>%
  count() %>%
  knitr::kable()
LevyByGroup %>%
  filter(CollabEfficiency<1) %>%
  count() %>%
  knitr::kable()

```


# Bayesian analysis

Here we look at (and compare) the proportions in the 4 Zipp groups.

Using model code for the Bayesian First Aid alternative to the test of proportions.
```{r bayes}
require(rjags)

source("DBDA2E-utilities.R")
source("DBDAderivatives.R")

myData = S123data %>%
  select(ZippGroup, Stage3score)

params = c(2,2)
# The model string written in the JAGS language
model_string <- paste0("model {
for(i in 1:length(x)) {
  x[i] ~ dbinom(theta[i], n[i])
  theta[i] ~ dbeta(",params[1],", ",params[2],")
  x_pred[i] ~ dbinom(theta[i], n[i])
}
}")

# Running the model
modelS3 <- jags.model(textConnection(model_string), data = list(x = zipptab$numcorrect, n = zipptab$numingroup), 
                    n.chains = 3, n.adapt=1000)
samplesS3 <- coda.samples(modelS3, c("theta", "x_pred"), n.iter=5000)

```

You can extract the mcmc samples as a matrix and compare the thetas 
of the groups. For example, the following shows the median and 95%
credible interval for the difference between Group 1 and Group 2.
```{r compare-thetas}
samp_mat <- as.matrix(samplesS3)
print(quantile(samp_mat[, "theta[2]"] - samp_mat[, "theta[1]"], c(0.025, 0.5, 0.975)))
print(quantile(samp_mat[, "theta[4]"] - samp_mat[, "theta[3]"], c(0.025, 0.5, 0.975)))

diagMCMC(samplesS3, parName = "theta[1]", saveName="Figs/Study2_S3props", saveType = "pdf")
diagMCMC(samplesS3, parName = "theta[2]", saveName="Figs/Study2_S3props", saveType = "pdf")
diagMCMC(samplesS3, parName = "theta[3]", saveName="Figs/Study2_S3props", saveType = "pdf")
diagMCMC(samplesS3, parName = "theta[4]", saveName="Figs/Study2_S3props", saveType = "pdf")

```

```{r mcmc-plot-s3props}
plotMCMC( samplesS3, data=myData, yName="Stage3score", sName="ZippGroup", compVal=NULL, compValDiff=0.0,
          saveName="Figs/Study2_S3props", saveType = "pdf")

contrasts = list(
  list( c(1,3), c(2,4), compVal=0.0, ROPE=c(-0.1,0.1)),
  list( c(1,2), c(3,4), compVal=0.0, ROPE=c(-0.1,0.1))
)
plotMCMCwithContrasts( samplesS3, datFrm=data.frame(myData), yName="Stage3score", xName="ZippGroup", contrasts = contrasts,
          saveName="Figs/Study2_S3props", saveType = "pdf")

```

