---
title: 'Two-stage exams: Study 3'
author: "George Kinnear"
date: "20/06/2020"
output:
  pdf_document:
    toc: true
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.path='rmd_figs_study3/',
                      warning=FALSE, message=FALSE)

library(tidyverse)
library(ggthemes)
library(ggpubr)
library(gt)        # For formatting tables


theme_set(
  theme_pubr() +
    theme(legend.position = "right")
)
heathers = c("#00b3c3","#e51c24","#016637","#fcb123")    #original green: #008a49

```

# Data

Import the dataset.

```{r data}
data = read.csv('Study3_data.csv', header=T) %>%
  dplyr::select(-c(X))


S123data = subset(data,
                  select=c('Q','Student','Stage1score','Stage2score','Stage3score')) %>%
  filter(!is.na(Stage2score))
```



## Mean at each stage (as bars)
```{r fig-means-bars}
ld <- gather(data = S123data, 
             key = stage, 
             value = score, 
             Stage1score, Stage2score, Stage3score)
ld <- ld[complete.cases(ld),]

plotData <- aggregate(ld$score,
                      by = list(Q = ld$Q, Stage = ld$stage),
                      FUN = function(x) c(mean = mean(x), sd = sd(x),
                                          n = length(x)))
plotData <- do.call(data.frame, plotData)
plotData$se <- plotData$x.sd / sqrt(plotData$x.n)

colnames(plotData) <- c("Q", "Stage", "mean", "sd", "n", "se")

plotData = plotData %>%
  mutate(
    Stage = gsub(".*(\\d).*","\\1",Stage)  # alternatively: Stage = parse_number(Stage)
  )
plotCounts = plotData %>% group_by(Q,Stage) %>% select(Stage,n) %>%
  mutate(scale_lab = paste0("Stage ",Stage, " (n=",n,")"))

limits <- aes(ymax = plotData$mean + plotData$se,
              ymin = plotData$mean - plotData$se)

p <- ggplot(data = plotData, aes(x = factor(Q), y = mean,
                                 fill = factor(Stage), label=n)) +
  geom_bar(stat = "identity",
             position = position_dodge(0.9))+
  geom_text(position = position_dodge(width = 0.9),aes(y=-0.05),angle=90) +
  geom_errorbar(limits, position = position_dodge(0.9),
                width = 0.25)  +
  labs(x = "Question",
       y = "Percentage of students answering correctly",
       fill = "Stage") +
  scale_fill_manual(values=heathers, labels=paste0("Stage ",c(1:4))) + #plotCounts$scale_lab) +
  #scale_fill_grey() +
  scale_y_continuous(labels = scales::percent)
p
ggsave("Figs/Study3_S123_means.pdf",width=20,height=10,units="cm",dpi=300)
```

A look at the data (this only shows the first few rows, but for a sanity check the full table could be consulted):
```{r data-peek}
S123data %>%
  group_by(Student) %>%
  mutate(
    Stage1sum = sum(Stage1score),
    Stage2sum = sum(Stage2score),
    Stage3sum = sum(Stage3score),
    qs = str_length(paste0(Stage1score, collapse=""))
  ) %>%
  ungroup() %>%
  mutate(
    S1max = max(Stage1sum)
  ) %>%
  arrange(-qs) %>%
  head() %>%
  knitr::kable()
```

## Bar plot
```{r fig-bars-expt}
barPlotData = data %>%
  dplyr::select(c('Q','Student','ZippGroup','Stage1score','Stage2score','Stage3score')) %>%
  gather('Stage1score','Stage2score','Stage3score',key="Stage", value="Score") %>%
  drop_na() %>%
  mutate(
    Stage=parse_number(Stage),
    expt = case_when(
      str_sub(ZippGroup,1,1)=="E" ~ "E",
      TRUE ~ "C"
    ),
    bar = case_when(
      Stage==1 & expt=="E" ~ "1E",
      Stage==1 & expt=="C" ~ "1C",
      Stage==2 ~ "2E",
      Stage==3 & expt=="E" ~ "3E",
      Stage==3 & expt=="C" ~ "3C"
    )
  ) %>%
  group_by(Q,bar) %>%
  summarise(
    mean=mean(Score,na.rm=TRUE),
    sd=sd(Score,na.rm=TRUE),
    n=n(),
    se=sd/sqrt(n)
  ) %>%
  ungroup() %>% 
  mutate(
    Q=paste0("Q",Q),
    Stage=parse_number(bar),
    Expt=str_sub(bar,2,2),
    ConditionOrder = case_when(
      bar=="1C" ~ 1,
      bar=="1E" ~ 2,
      bar=="2E" ~ 3,
      bar=="3E" ~ 4,
      bar=="3C" ~ 5
    ),
    bar2=fct_reorder(bar,ConditionOrder)
  ) %>% arrange(Q,ConditionOrder)


barLabels = c("Stage 1 (Control)", "Stage 1 (Experimental)",
              "Stage 2 (Experimental)",
              "Stage 3 (Experimental)", "Stage 3 (Control)")
barColoursAlpha = c(alpha(heathers[1],.25),heathers[1],heathers[2],heathers[3],alpha(heathers[3],.25))
barColoursAlpha = c("#bfecf0",heathers[1],heathers[2],heathers[3],"#bfd9cd")

limits <- aes(ymax = barPlotData$mean + barPlotData$se,
              ymin = barPlotData$mean - barPlotData$se)

ggplot(data = barPlotData, aes(x = factor(Q), y = mean,
                                    fill = factor(bar), label=n)) +
  geom_bar(stat = "identity",
           position = position_dodge(0.9),color="white")+
  geom_text(position = position_dodge(width = 0.9),aes(y=-0.05),angle=90) +
  geom_errorbar(limits, position = position_dodge(0.9),
                width = 0.25)  +
  labs(x = "Question",
       y = "Percentage of students answering correctly",
       fill = "Stage") +
  scale_fill_manual(values=barColoursAlpha, labels=barLabels) +
  scale_y_continuous(labels = scales::percent)
ggsave("Figs/Study3_S123_means.pdf",width=20,height=10,units="cm",dpi=300)
```


## Mean at each stage (as points)
```{r fig-expt-pts}

ggplot(data=barPlotData,aes(x=bar2,y=mean,group=Q,label=n))+
  geom_line()+
  geom_errorbar(aes(ymax = barPlotData$mean + barPlotData$se,
                    ymin = barPlotData$mean - barPlotData$se),
                position = position_dodge(0.9),
                width = 0.5)  +
  geom_point(position = position_dodge(0.9),size=5,
             aes(color=bar2))+
  facet_grid(cols=vars(Q))+
  labs(x = "Stage",
       y = "Percentage of students answering correctly",
       color = "Stage") +
  scale_y_continuous(labels = scales::percent)+
  scale_color_manual(values=barColoursAlpha, labels=barLabels)+
  coord_cartesian(ylim=c(0,1),clip="off")+
  geom_text(position = position_dodge(width = 0.9),
            aes(y=0, label=paste0("",barPlotData$n)),
            angle=90,
            color="#777777") +
  theme(strip.background = element_rect(fill=NA,colour = NA),
        strip.text = element_text(size=12, face="bold"),
        axis.text.x = element_text(angle=90))
ggsave("Figs/Study3_S123_means_pts.pdf",width=20,height=10,units="cm",dpi=300)

```

## Mean at each stage (table, with standard errors)

```{r tab-means}
tab = barPlotData %>% 
  mutate(
    entry = paste0(sprintf("%2.0f", mean*100), " (", sprintf("%2.1f", se*100), ")")
  ) %>%
  group_by(Q,bar) %>%
  select(Q,bar,entry) %>%
  spread(Q,entry)

tab$bar = barLabels
tab %>% knitr::kable()
```


# Forming the Zipp tables

This constructs the data in Table 5 of the paper. Note that this only has data for the experimental condition -- the data for the full Table 5 (i.e. including the control condition) appears in the final section of this script, where the experimental analysis takes place.

## Table 5, first attempt
```{r zip-tab-123}
S123data = subset(data,select=c('Q','Student','Stage1score','Stage2score','Stage3score','ZippGroup'))
S123data = S123data[complete.cases(S123data),]

zipptab = S123data %>%
  group_by(ZippGroup) %>%
  summarize( numcorrect = sum(Stage3score),
             numingroup = n()) %>%
  mutate( pc = numcorrect/numingroup,
          entry = paste0(sprintf("%2.1f", pc*100), " (", numcorrect, "/",numingroup,")"))
zipptab %>% knitr::kable()
```


# Group dynamics: Stage 1 vs Stage 2

Here we look at the relative performance in the groups across the first two stages.

```{r group-correctness}

groupCorrectness = data %>%
  group_by(Stage2group,Q) %>%
  summarise(
    GpSize = n(),
    S1sum = sum(Stage1score),
    S1avg = S1sum/GpSize,
    S2 = max(Stage2score),
    S2pc = ceiling(max(Stage2scorePC)) ## round up so that it's 1 if they were correct on second attempt
  ) %>%
  filter(
    !is.na(S2)
  )
groupCorrectness %>% ungroup() %>% knitr::kable()
```

```{r group-perfs}
groupPerfS12 = groupCorrectness %>%
  mutate(
    tot_group = cut(S1sum,breaks=c(-Inf,0.5,1.5,2.5,Inf),labels=c("0","1","2","3 or more"))
  ) %>%
  group_by(tot_group) %>%
  summarize(
    S2avg = mean(S2),
    S2se = sd(S2)/sqrt(n()),
    S2n = n(),
    S2Pavg = mean(S2pc),
    S2Pse = sd(S2pc)/sqrt(n())
  )
groupPerfS12 %>% knitr::kable()
```

```{r group-perf-plot}

ggplot(groupPerfS12,aes(x=tot_group,y=S2avg,label=S2n))+
  geom_errorbar(aes(ymax = groupPerfS12$S2avg + groupPerfS12$S2se,
                    ymin = groupPerfS12$S2avg - groupPerfS12$S2se),
                position = position_dodge(0.9),
                width = 0.1)+
  geom_point(aes(colour="First attempt"),size=5)+
  geom_errorbar(aes(ymax = groupPerfS12$S2Pavg + groupPerfS12$S2Pse,
                    ymin = groupPerfS12$S2Pavg - groupPerfS12$S2Pse),
                position = position_nudge(x=0.1),
                width = 0.1)+
  geom_point(aes(y=S2Pavg,colour="Second attempt"),position=position_nudge(x=0.1),size=5)+
  scale_y_continuous(labels = scales::percent,breaks=seq(0,1,by=.2))+
  scale_color_manual(values=heathers) +
  coord_cartesian(ylim=c(0,1),clip="off")+
  geom_text(position = position_dodge(width = 0.9),
            aes(y=-0.01, label=paste0("n=",groupPerfS12$S2n)),
            angle=0,
            color="#777777")+
  labs(x = "Number of students correct at stage 1",
       y = "Groups answering correctly",
       colour = "Stage 2 attempt") +
  theme(strip.background = element_rect(fill=NA,colour = NA),
        strip.text = element_text(size=12, face="bold"))
ggsave("Figs/Study3_S12_collab.pdf",width=15,height=7,units="cm",dpi=300)

```


## Group dynamics

This replicates the analysis of Levy et al. (2018), producing Fig 7 of the paper. There is extra detail here, with the various measures like `collaborative efficiency' shown for each group and also plotted.

Find the top scoring student in each group,
and the "super" score (max score across all students in the group, by question)


```{r analyse-stages12}

S12data_scored = data %>%
  dplyr::select(Q,Stage1score,Stage2score,Student,Stage2group) %>%
  mutate(
    Group = Stage2group
  ) %>%
  dplyr::select(-Stage2group)

S1superandtop = S12data_scored %>%
  group_by(Group,Q) %>%
  mutate(
    superstudent = max(Stage1score)
  ) %>%
  group_by(Group,Student) %>%
  mutate(
    topstudent = sum(Stage1score)/n() # the Student's mean score on the n() Questions
  ) %>%
  group_by(Group) %>%
  summarise(
    superstudent = sum(superstudent)/n(),
    topstudent = max(topstudent)
  )


LevyA = S12data_scored %>%
  group_by(Student) %>%
  summarise(
    Stage1pc = sum(Stage1score)/n()
  ) %>%
  summarise(
    S1mean = mean(Stage1pc),
    S1sd = sd(Stage1pc),
    S1n = n()
  )


LevyAsd =  LevyA$S1sd[[1]] 

groupCorrectness = data %>%
  group_by(Stage2group,Q) %>%
  summarise(
    GpSize = n(),
    S1sum = sum(Stage1score),
    S1avg = S1sum/GpSize,
    S2 = max(Stage2score)
  )

LevyByGroup = groupCorrectness %>%
  mutate(
    Group = Stage2group
  ) %>%
  left_join(S1superandtop) %>%
#  left_join(LevyA %>% select(S1sd)) %>%
  group_by(Group) %>%
  summarise(
    n = max(GpSize),
    IndivA = mean(S1avg),
    GroupB = mean(S2,na.rm=TRUE),
    TopC = max(topstudent),
    SuperD = max(superstudent),
    GainBA = (GroupB-IndivA)/LevyAsd,
    TopSurplus = (TopC-IndivA)/LevyAsd,
    SuperSurplus = (SuperD-IndivA)/LevyAsd,
    CollabEfficiency = GainBA / na_if(SuperSurplus,0)
  ) %>%
  ungroup()

LevyByGroup %>% knitr::kable(digits = 2)
```

```{r levy-plots}
ggplot(stack(LevyByGroup %>% select(IndivA,GroupB,TopC,SuperD)), aes(x = ind, y = values)) +
  geom_violin(fill=heathers[1]) +
  geom_boxplot(width=0.2,color=heathers[2],lwd=1) +
  geom_jitter(shape=16, position=position_jitter(0.2),alpha=0.5) +
  labs(x = "Average score of...",
       y = "Percentage correct") +
  scale_y_continuous(labels = scales::percent)
ggsave("Figs/Study3_LevyABCD.pdf",width=20,height=10,units="cm",dpi=300)
ggsave("Figs/Study3_LevyABCD_small.pdf",width=10,height=7,units="cm",dpi=300)

LevyByGroup %>% select(GainBA,TopSurplus,SuperSurplus,CollabEfficiency) %>%
  filter(CollabEfficiency>4)


ggplot(stack(LevyByGroup %>% select(GainBA,TopSurplus,SuperSurplus,CollabEfficiency) %>%
               mutate(CollabEfficiency = ifelse(CollabEfficiency<4,CollabEfficiency,NA_real_))), aes(x = ind, y = values)) +
  geom_violin(fill=heathers[1]) +
  geom_boxplot(width=0.2,color=heathers[2],lwd=1) +
  geom_jitter(shape=16, position=position_jitter(0.2),alpha=0.5) + 
  labs(x = "Difference in average scores",
       y = "Difference (in SDs)")+
  theme(axis.text.x = element_text(angle = 15, hjust = 1))
ggsave("Figs/Study3_LevyDiffs.pdf",width=20,height=10,units="cm",dpi=300)
ggsave("Figs/Study3_LevyDiffs_small.pdf",width=10,height=7,units="cm",dpi=300)


LevyByGroup %>%
  summarise(
    CollabEfficiency_m = mean(CollabEfficiency, na.rm=TRUE),
    CollabEfficiency_sd = sd(CollabEfficiency, na.rm=TRUE),
    n=n()
  ) %>% knitr::kable(digits = 2)

LevyByGroup %>%
  filter(CollabEfficiency>1) %>%
  count()

```


# Bayesian analysis

Here we look at (and compare) the proportions in the 4 experimental groups shown in Table 5 of the paper.

Using model code for the Bayesian First Aid alternative to the test of proportions.
```{r bayes}
require(rjags)

source("DBDA2E-utilities.R")
source("DBDAderivatives.R")

myData = S123data %>%
  select(ZippGroup, Stage3score)


myData = S123data %>%
  select(ZippGroup, Stage3score) %>%
  mutate(
    ZippGroup = as.numeric(str_sub(ZippGroup,2,2))
  )

params = c(2,2)
# The model string written in the JAGS language
model_string <- paste0("model {
for(i in 1:length(x)) {
  x[i] ~ dbinom(theta[i], n[i])
  theta[i] ~ dbeta(",params[1],", ",params[2],")
  x_pred[i] ~ dbinom(theta[i], n[i])
}
}")

# Running the model
modelS3 <- jags.model(textConnection(model_string), data = list(x = zipptab$numcorrect, n = zipptab$numingroup), 
                    n.chains = 3, n.adapt=1000)
samplesS3 <- coda.samples(modelS3, c("theta", "x_pred"), n.iter=5000)

```

You can extract the mcmc samples as a matrix and compare the thetas 
of the groups. For example, the following shows the median and 95%
credible interval for the difference between Group 1 and Group 2.
```{r compare-thetas}
samp_mat <- as.matrix(samplesS3)
print(quantile(samp_mat[, "theta[2]"] - samp_mat[, "theta[1]"], c(0.025, 0.5, 0.975)))
print(quantile(samp_mat[, "theta[4]"] - samp_mat[, "theta[3]"], c(0.025, 0.5, 0.975)))

diagMCMC(samplesS3, parName = "theta[1]", saveName="Figs/Study3_S3props", saveType = "pdf")
diagMCMC(samplesS3, parName = "theta[2]", saveName="Figs/Study3_S3props", saveType = "pdf")
diagMCMC(samplesS3, parName = "theta[3]", saveName="Figs/Study3_S3props", saveType = "pdf")
diagMCMC(samplesS3, parName = "theta[4]", saveName="Figs/Study3_S3props", saveType = "pdf")


```

```{r mcmc-plot-s3props}
plotMCMC( samplesS3, data=myData, yName="Stage3score", sName="ZippGroup", compVal=NULL, compValDiff=0.0,
          saveName="Figs/Study3_S3props", saveType = "pdf")

contrasts = list(
  list( c(1,3), c(2,4), compVal=0.0, ROPE=c(-0.1,0.1)),
  list( c(1,2), c(3,4), compVal=0.0, ROPE=c(-0.1,0.1))
)
plotMCMCwithContrasts( samplesS3, datFrm=data.frame(myData), yName="Stage3score", xName="ZippGroup", contrasts = contrasts,
          saveName="Figs/Study3_S3props", saveType = "pdf")


```

# Experimental analysis

This conducts the Bayesian analysis of the main experiment, i.e. comparing all 6 groups from Table 5.

```{r expt-analysis-dataprep}

myData = data %>%
  dplyr::select(ZippGroup, Stage3score) %>%
  drop_na() %>%
  mutate(ZippGroup = fct_relevel(ZippGroup, "C0", after=Inf)) %>%
  mutate(ZippGroup = fct_relevel(ZippGroup, "C1", after=Inf)) %>%
  mutate(
    ZippGroup = case_when(
      ZippGroup=="C0" ~ 5,
      ZippGroup=="C1" ~ 6,
      TRUE ~ as.numeric(str_sub(ZippGroup,2,2))
    )
  )

myData %>% group_by(ZippGroup) %>% summarise( mean(Stage3score)) %>% knitr::kable()
```

```{r zipp-tab-table5}
zipptab = myData %>%
  group_by(ZippGroup) %>%
  summarise(
    numcorrect = sum(Stage3score),
    numingroup = n()
  )
zipptab %>% knitr::kable()
```

```{r expt-analysis}
params = c(2,2)
# The model string written in the JAGS language
model_string <- paste0("model {
                       for(i in 1:length(x)) {
                       x[i] ~ dbinom(theta[i], n[i])
                       theta[i] ~ dbeta(",params[1],", ",params[2],")
                       x_pred[i] ~ dbinom(theta[i], n[i])
                       }
                       }")

# Running the model
modelEXPT <- jags.model(textConnection(model_string), data = list(x = zipptab$numcorrect, n = zipptab$numingroup), 
                      n.chains = 3, n.adapt=1000)
samplesEXPT <- coda.samples(modelEXPT, c("theta", "x_pred"), n.iter=5000)

# Inspecting the posterior
#plot(samples)
#summary(samples)

# You can extract the mcmc samples as a matrix and compare the thetas 
# of the groups. For example, the following shows the median and 95%
# credible interval for the difference between Group 1 and Group 2.
samp_mat <- as.matrix(samplesEXPT)
print(quantile(samp_mat[, "theta[2]"] - samp_mat[, "theta[1]"], c(0.025, 0.5, 0.975)))
print(quantile(samp_mat[, "theta[4]"] - samp_mat[, "theta[3]"], c(0.025, 0.5, 0.975)))

diagMCMC(samplesEXPT, parName = "theta[1]", saveName="Figs/Study3_EXPTprops", saveType = "pdf")
diagMCMC(samplesEXPT, parName = "theta[2]", saveName="Figs/Study3_EXPTprops", saveType = "pdf")
diagMCMC(samplesEXPT, parName = "theta[3]", saveName="Figs/Study3_EXPTprops", saveType = "pdf")
diagMCMC(samplesEXPT, parName = "theta[4]", saveName="Figs/Study3_EXPTprops", saveType = "pdf")
diagMCMC(samplesEXPT, parName = "theta[5]", saveName="Figs/Study3_EXPTprops", saveType = "pdf")
diagMCMC(samplesEXPT, parName = "theta[6]", saveName="Figs/Study3_EXPTprops", saveType = "pdf")

#plotPost( samplesS3[,"theta[1]"] , main="theta[1]" , xlab=bquote(theta[1]) )
#plotPost( samplesS3[,"theta[2]"] , main="theta[2]" , xlab=bquote(theta[2]) )
#plotPost( samplesS3[,"theta[3]"] , main="theta[3]" , xlab=bquote(theta[3]) )
#plotPost( samplesS3[,"theta[4]"] , main="theta[4]" , xlab=bquote(theta[4]) )

plotMCMC( samplesEXPT, data=myData, yName="Stage3score", sName="ZippGroup", compVal=NULL, compValDiff=0.0,
          saveName="Figs/Study3_EXPTprops", saveType = "pdf")

contrasts = list(
  list( c(1,3), c(2,4), compVal=0.0, ROPE=c(-0.1,0.1)),
  list( c(1,2), c(3,4), compVal=0.0, ROPE=c(-0.1,0.1)),
  list( c(1,2,3,4), c(5,6), compVal=0.0, ROPE=c(-0.1,0.1)),
  list( c(1,2), c(5), compVal=0.0, ROPE=c(-0.1,0.1)),
  list( c(3,4), c(6), compVal=0.0, ROPE=c(-0.1,0.1))
)

plotMCMCwithContrasts( samplesEXPT, datFrm=data.frame(myData), yName="Stage3score", xName="ZippGroup", contrasts = contrasts,
                       saveName="Figs/Study3_EXPTprops", saveType = "pdf")

```