---
title: 'Two-stage exams: Study 1'
author: "George Kinnear"
date: "20/06/2020"
output:
  pdf_document:
    toc: true
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.path='rmd_figs_study1/',
                      warning=FALSE, message=FALSE)

library(tidyverse)
library(ggthemes)
library(ggpubr)
library(gt)        # For formatting tables


theme_set(
  theme_pubr() +
    theme(legend.position = "right")
)
heathers = c("#00b3c3","#e51c24","#016637","#fcb123")    #original green: #008a49

```

# Data

Import and combine the datasets.

```{r data}

data2017 = read.csv('study1_2017.csv', header=T)
data2018 = read.csv('study1_2018.csv', header=T)

data1718 = bind_rows("2017"=data2017, "2018"=data2018, .id="Year") %>%
  select(-c(X)) %>%
  mutate(Q = fct_relevel(Q, c("Q1a", "Q1b", "Q2a", "Q2b", "Q2c", "Q11")))


S1234data = subset(data1718,
                   select=c('Q','Student','Stage1score','Stage2score','Stage3score','Stage4score')) %>%
  filter(!is.na(Stage2score))


S123data = subset(data1718,select=c('Q','Student','Stage1score','Stage2score','Stage3score','ZippGroup'))
S123data = S123data[complete.cases(S123data),]
```


## Check that it makes sense to combine the two datasets

Viewing both sets of data to check that the items performed similarly in both years.

```{r combine-data-check}
ld <- gather(data = subset(data1718,select=c('Year','Q','Student','Stage1score','Stage2score','Stage3score')), 
             key = stage, 
             value = score, 
             Stage1score, Stage2score, Stage3score)
ld <- ld[complete.cases(ld),]

# Code from here:
# https://www.r-bloggers.com/building-barplots-with-error-bars/

plotData <- aggregate(ld$score,
                      by = list(Year = ld$Year, Q = ld$Q, Stage = ld$stage),
                      FUN = function(x) c(mean = mean(x), sd = sd(x),
                                          n = length(x)))
plotData <- do.call(data.frame, plotData)
plotData$se <- plotData$x.sd / sqrt(plotData$x.n)

colnames(plotData) <- c("Year","Q", "Stage", "mean", "sd", "n", "se")

limits <- aes(ymax = plotData$mean + plotData$se,
              ymin = plotData$mean - plotData$se)

p <- ggplot(data = plotData, aes(x = factor(Q), y = mean,
                                 fill = factor(Stage))) +
  facet_grid(Year ~ .)+
  geom_bar(stat = "identity",
           position = position_dodge(0.9)) +
  labs(x = "Question",
       y = "Percentage of students answering correctly",
       fill = "Stage") +
  scale_fill_manual(values=heathers) + #viridis_pal(1,0,1)(4)) +
  scale_y_continuous(labels = scales::percent)
p
```


## Mean at each stage (as bars)
```{r fig-means-bars}
ld <- gather(data = S1234data, 
             key = stage, 
             value = score, 
             Stage1score, Stage2score, Stage3score, Stage4score)
ld <- ld[complete.cases(ld),]

plotData <- aggregate(ld$score,
                      by = list(Q = ld$Q, Stage = ld$stage),
                      FUN = function(x) c(mean = mean(x), sd = sd(x),
                                          n = length(x)))
plotData <- do.call(data.frame, plotData)
plotData$se <- plotData$x.sd / sqrt(plotData$x.n)

colnames(plotData) <- c("Q", "Stage", "mean", "sd", "n", "se")

plotData = plotData %>%
  mutate(
    Q = fct_relevel(Q, c( "Q1b", "Q2a", "Q2b","Q1a","Q2c",  "Q11")),
    Stage = gsub(".*(\\d).*","\\1",Stage)  # alternatively: Stage = parse_number(Stage)
  )
plotCounts = plotData %>% filter(Q=="Q1b") %>% group_by(Stage) %>% select(Stage,n) %>%
  mutate(scale_lab = paste0("Stage ",Stage, " (n=",n,")"))

limits <- aes(ymax = plotData$mean + plotData$se,
              ymin = plotData$mean - plotData$se)

p <- ggplot(data = plotData, aes(x = factor(Q), y = mean,
                                 fill = factor(Stage), label=n)) +
  geom_bar(stat = "identity",
             position = position_dodge(0.9))+
  geom_errorbar(limits, position = position_dodge(0.9),
                width = 0.25)  +
  labs(x = "Question",
       y = "Percentage of students answering correctly",
       fill = "Stage") +
  scale_fill_manual(values=heathers, labels=plotCounts$scale_lab) +
  scale_y_continuous(labels = scales::percent)
ggsave("Figs/Study1_S1234_means.pdf",width=20,height=10,units="cm",dpi=300)
```

## Mean at each stage (as points)
```{r fig-means-pts}
ggplot(data=plotData,aes(x=Stage,y=mean,group=Q))+
  geom_line()+
  geom_errorbar(limits, position = position_dodge(0.9),
                width = 0.5)  +
  geom_point(position = position_dodge(0.9),size=5,aes(color=Stage))+
  scale_color_manual(values=heathers, labels=plotCounts$scale_lab) +
  labs(x = "Stage",
       y = "Percentage of students answering correctly",
       fill = "Stage") +
  scale_y_continuous(labels = scales::percent, limits=c(0,1))+
  facet_grid(cols=vars(Q))+
  theme(strip.background = element_rect(fill=NA,colour = NA),
        strip.text = element_text(size=12, face="bold"))
ggsave("Figs/Study1_S1234_means_pts.pdf",width=20,height=10,units="cm",dpi=300)
```

## Mean at each stage (table, with standard errors)

```{r tab-means}
tab = plotData %>% 
  mutate(
    entry = paste0(sprintf("%2.0f", mean*100), " (", sprintf("%2.1f", se*100), ")"),
    Stage = gsub(".*(\\d).*","\\1",Stage)
  ) %>%
  group_by(Q,Stage) %>%
  select(Q,Stage,entry) %>%
  spread(Q,entry)

tab %>% knitr::kable()
#tab %>% knitr::kable(format="latex",booktabs=T)
```


# Forming the Zipp tables

This constructs the data in Table 3 of the paper.

## Stages 1-3 only
```{r zip-tab-123}
zipptab = S123data %>%
  group_by(ZippGroup) %>%
  summarize( numcorrect = sum(Stage3score),
             numingroup = n()) %>%
  mutate( pc = numcorrect/numingroup,
          entry = paste0(sprintf("%2.1f", pc*100), " (", numcorrect, "/",numingroup,")"))
zipptab %>% knitr::kable()
```

## Stages 1-4
```{r zip-tab-1234}
S124data = subset(data1718,select=c('Q','Student','Stage1score','Stage2score','Stage4score','ZippGroup'))
S124data = S124data[complete.cases(S124data),]

zipptab124 = S124data %>%
  group_by(ZippGroup) %>%
  summarize( numcorrect = sum(Stage4score),
             numingroup = n()) %>%
  mutate( pc = numcorrect/numingroup,
          entry = paste0(sprintf("%2.1f", pc*100), " (", numcorrect, "/",numingroup,")"))
zipptab124 %>% knitr::kable()
```


# Group dynamics: Stage 1 vs Stage 2

Here we look at the relative performance in the groups across the first two stages.

```{r group-correctness}

groupCorrectness = data1718 %>%
  mutate(
    Stage2group = paste0(Year,"_",Stage2group)
  ) %>%
  group_by(Stage2group,Q) %>%
  summarise(
    GpSize = n(),
    S1sum = sum(Stage1score),
    S1avg = S1sum/GpSize,
    S2 = max(Stage2score)
  )
groupCorrectness %>% ungroup() %>% gt()
```

```{r group-perfs}
groupPerfS12 = groupCorrectness %>%
  mutate(
    tot_group = cut(S1sum,breaks=c(-Inf,0.5,1.5,2.5,Inf),labels=c("0","1","2","3 or more"))
  ) %>%
  group_by(tot_group) %>%
  summarize(
    S2avg = mean(S2),
    S2se = sd(S2)/sqrt(n()),
    S2n = n()
  )
groupPerfS12 %>% knitr::kable()
```

```{r group-perf-plot}
ggplot(groupPerfS12,aes(x=tot_group,y=S2avg,label=S2n))+
  geom_errorbar(aes(ymax = groupPerfS12$S2avg + groupPerfS12$S2se,
                    ymin = groupPerfS12$S2avg - groupPerfS12$S2se),
                position = position_dodge(0.9),
                width = 0.1)+
  geom_point(size=5,colour=heathers[1])+
  scale_y_continuous(labels = scales::percent,breaks=seq(0,1,by=.2))+
  scale_color_manual(values=heathers) +
  coord_cartesian(ylim=c(0,1),clip="off")+
  geom_text(position = position_dodge(width = 0.9),
            aes(y=-0.01, label=paste0("n=",groupPerfS12$S2n)),
            angle=0,
            color="#777777")+
  labs(x = "Number of students correct at stage 1",
       y = "Groups answering correctly",
       colour = "Stage 2 attempt") +
  theme(strip.background = element_rect(fill=NA,colour = NA),
        strip.text = element_text(size=12, face="bold"))
ggsave("Figs/Study1_S12_collab.pdf",width=15,height=7,units="cm",dpi=300)

```


## Group dynamics

This replicates the analysis of Levy et al. (2018), producing Fig 2 of the paper. There is extra detail here, with the various measures like `collaborative efficiency' shown for each group and also plotted.

```{r analyse-stages12}

S12data_scored = data1718 %>%
  dplyr::select(Year,Q,Stage1score,Stage2score,Student,Stage2group) %>%
  mutate(
    Group = paste0(Year,"_",Stage2group)
  ) %>%
  dplyr::select(-Stage2group)

S1superandtop = S12data_scored %>%
  group_by(Group,Q) %>%
  mutate(
    superstudent = max(Stage1score)
  ) %>%
  group_by(Group,Student) %>%
  mutate(
    topstudent = sum(Stage1score)/n() # the Student's mean score on the n() Questions
  ) %>%
  group_by(Group) %>%
  summarise(
    superstudent = sum(superstudent)/n(),
    topstudent = max(topstudent)
  )

LevyA = S12data_scored %>%
  group_by(Student) %>%
  summarise(
    Stage1pc = sum(Stage1score)/n()
  ) %>%
  summarise(
    S1mean = mean(Stage1pc),
    S1sd = sd(Stage1pc),
    S1n = n()
  )

LevyAsd =  LevyA$S1sd[[1]] 

LevyByGroup = groupCorrectness %>%
  mutate(
    Group = Stage2group
  ) %>%
  left_join(S1superandtop) %>%
  group_by(Group) %>%
  summarise(
    n = max(GpSize),
    IndivA = mean(S1avg),
    GroupB = mean(S2),
    TopC = max(topstudent),
    SuperD = max(superstudent),
    GainBA = (GroupB-IndivA)/LevyAsd,
    TopSurplus = (TopC-IndivA)/LevyAsd,
    SuperSurplus = (SuperD-IndivA)/LevyAsd,
    CollabEfficiency = GainBA / SuperSurplus
  ) %>%
  ungroup()
LevyByGroup %>% knitr::kable()
```

```{r levy-plots}
ggplot(stack(LevyByGroup %>% select(IndivA,GroupB,TopC,SuperD)), aes(x = ind, y = values)) +
  geom_violin(fill=heathers[1]) +
  geom_boxplot(width=0.2,color=heathers[2],lwd=1) +
  geom_jitter(shape=16, position=position_jitter(0.2)) +
  labs(x = "Average score of...",
       y = "Percentage correct") +
  scale_y_continuous(labels = scales::percent)
ggsave("Figs/Study1_LevyABCD.pdf",width=20,height=10,units="cm",dpi=300)
ggsave("Figs/Study1_LevyABCD_small.pdf",width=10,height=7,units="cm",dpi=300)

ggplot(stack(LevyByGroup %>% select(GainBA,TopSurplus,SuperSurplus,CollabEfficiency)), aes(x = ind, y = values)) +
  geom_violin(fill=heathers[1]) +
  geom_boxplot(width=0.2,color=heathers[2],lwd=1) +
  geom_jitter(shape=16, position=position_jitter(0.2)) + 
  labs(x = "Difference in average scores",
       y = "Difference (in SDs)")+
  theme(axis.text.x = element_text(angle = 15, hjust = 1))
ggsave("Figs/Study1_LevyDiffs.pdf",width=20,height=10,units="cm",dpi=300)
ggsave("Figs/Study1_LevyDiffs_small.pdf",width=10,height=7,units="cm",dpi=300)
```

```{r levy-summary}
LevyByGroup %>%
  summarise(
    CollabEfficiency_m = mean(CollabEfficiency),
    CollabEfficiency_sd = sd(CollabEfficiency)
  ) %>% knitr::kable()
```


# Bayesian analysis

Here we look at (and compare) the proportions in the 4 Zipp groups.

Using model code for the Bayesian First Aid alternative to the test of proportions.
```{r bayes}
require(rjags)

source("DBDA2E-utilities.R")
source("DBDAderivatives.R")

myData = S123data %>%
  select(ZippGroup, Stage3score)

params = c(2,2)
# The model string written in the JAGS language
model_string <- paste0("model {
for(i in 1:length(x)) {
  x[i] ~ dbinom(theta[i], n[i])
  theta[i] ~ dbeta(",params[1],", ",params[2],")
  x_pred[i] ~ dbinom(theta[i], n[i])
}
}")

# Running the model
modelS3 <- jags.model(textConnection(model_string), data = list(x = zipptab$numcorrect, n = zipptab$numingroup), 
                    n.chains = 3, n.adapt=1000)
samplesS3 <- coda.samples(modelS3, c("theta", "x_pred"), n.iter=5000)

# Inspecting the posterior
#plot(samples)
#summary(samples)
```

You can extract the mcmc samples as a matrix and compare the thetas 
of the groups. For example, the following shows the median and 95%
credible interval for the difference between Group 1 and Group 2.
```{r compare-thetas}
samp_mat <- as.matrix(samplesS3)
print(quantile(samp_mat[, "theta[2]"] - samp_mat[, "theta[1]"], c(0.025, 0.5, 0.975)))
print(quantile(samp_mat[, "theta[4]"] - samp_mat[, "theta[3]"], c(0.025, 0.5, 0.975)))
```
```{r mcmc-plot-thetas}
diagMCMC(samplesS3, parName = "theta[1]", saveName="Figs/Study1_S3props", saveType = "pdf")
diagMCMC(samplesS3, parName = "theta[2]", saveName="Figs/Study1_S3props", saveType = "pdf")
diagMCMC(samplesS3, parName = "theta[3]", saveName="Figs/Study1_S3props", saveType = "pdf")
diagMCMC(samplesS3, parName = "theta[4]", saveName="Figs/Study1_S3props", saveType = "pdf")
```

```{r mcmc-plot-s3props}
plotMCMC( samplesS3, data=myData, yName="Stage3score", sName="ZippGroup", compVal=NULL, compValDiff=0.0,
          saveName="Figs/Study1_S3props", saveType = "pdf")

contrasts = list(
  list( c(1,3), c(2,4), compVal=0.0, ROPE=c(-0.1,0.1)),
  list( c(1,2), c(3,4), compVal=0.0, ROPE=c(-0.1,0.1))
)
plotMCMCwithContrasts( samplesS3, datFrm=data.frame(myData), yName="Stage3score", xName="ZippGroup", contrasts = contrasts,
          saveName="Figs/Study1_S3props", saveType = "pdf")
```


## For Stage 4

```{r bayes-stage4}
myData124 = S124data %>%
  select(ZippGroup, Stage4score)

params = c(2,2)
# The model string written in the JAGS language
model_string <- paste0("model {
 for(i in 1:length(x)) {
   x[i] ~ dbinom(theta[i], n[i])
   theta[i] ~ dbeta(",params[1],", ",params[2],")
   x_pred[i] ~ dbinom(theta[i], n[i])
 }
 }")

# Running the model
modelS4 <- jags.model(textConnection(model_string), data = list(x = zipptab124$numcorrect, n = zipptab124$numingroup), 
                      n.chains = 3, n.adapt=1000)
samplesS4 <- coda.samples(modelS4, c("theta", "x_pred"), n.iter=5000)

samp_mat <- as.matrix(samplesS4)
print(quantile(samp_mat[, "theta[2]"] - samp_mat[, "theta[1]"], c(0.025, 0.5, 0.975)))
print(quantile(samp_mat[, "theta[4]"] - samp_mat[, "theta[3]"], c(0.025, 0.5, 0.975)))
```

```{r mcmc-stage4-thetas}
diagMCMC(samplesS4, parName = "theta[1]", saveName="Figs/Study1_S4props", saveType = "pdf")
diagMCMC(samplesS4, parName = "theta[2]", saveName="Figs/Study1_S4props", saveType = "pdf")
diagMCMC(samplesS4, parName = "theta[3]", saveName="Figs/Study1_S4props", saveType = "pdf")
diagMCMC(samplesS4, parName = "theta[4]", saveName="Figs/Study1_S4props", saveType = "pdf")
```

```{r mcmc-stage4-contrasts}
plotMCMC( samplesS4, data=myData124, yName="Stage4score", sName="ZippGroup", compVal=NULL, compValDiff=0.0,
          saveName="Figs/Study1_S4props", saveType = "pdf")

contrasts = list(
  list( c(1,3), c(2,4), compVal=0.0, ROPE=c(-0.1,0.1)),
  list( c(1,2), c(3,4), compVal=0.0, ROPE=c(-0.1,0.1))
)
plotMCMCwithContrasts( samplesS4, datFrm=data.frame(myData124), yName="Stage4score", xName="ZippGroup", contrasts = contrasts,
                       saveName="Figs/Study1_S4props", saveType = "pdf")

```
